---
title: methods
bibliography: germanic-ai.bib
---

## methods
### snc
16062.1

### data

```{r}
#| echo: false
#| output: false

mh<-lmdf.c$target=="0-human"
mp<-lmdf.c$target=="post"
mg<-lmdf.c$target=="gpt"

sh<-sum(lmdf.c$freq[c(which(mh))])
sp<-sum(lmdf.c$freq[c(which(mp))])
sg<-sum(lmdf.c$freq[c(which(mg))])
ts<-data.frame(target=c("gemini","human-pre","human-post"),tokens=c(sg,sh,sp))

```

our human langugae data consists of raw texts from german bundestag plenary protocols (@dip_dip_2026). the LLM corpus consists of  model summaries of a first subset of these texts generated with the following prompt: @sec-gemini-prompt.

#### corpus subsets
`r knitr::kable(ts)`

### gemini prompt {#sec-gemini-prompt}
```{r}
#| echo: false
#| warning: false

ptx<-readLines("../g-prompt01.txt")
print(ptx)
```

### computation
we first devised AI-typical lemma in the model corpus which are distinctive for that corpus using a linear model (R, lme4::glmer()) that calculates a score for each lemma in the corpus 
