% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Noto Serif}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\usepackage{polyglossia}
\setmainlanguage{english}
\setotherlanguage{hebrew}
\newfontfamily\hebrewfont{Noto Serif Hebrew}[Script=Hebrew]
\newcommand{\he}[1]{\bgroup\textdir TRT{\hebrewfont #1}\egroup}
\KOMAoption{captions}{tableheading}
\usepackage{fancyvrb}
\usepackage{fvextra}
\RecustomVerbatimEnvironment{verbatim}{Verbatim}{breaklines=true,breakanywhere=true,fontsize=\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={balkons \& die straszenbahn},
  pdfauthor={st. schwarz},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{balkons \& die straszenbahn}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{cross language observations on The Shibboleth in yiddish,
frisian and berlin dialect}
\author{st. schwarz}
\date{2025-11-15}
\begin{document}
\maketitle


\section{index}\label{index}

\subsection{abstract A: idee 1}\label{abstract-a-idee-1}

In this paper I want to explore shibboleth phenomena in germanic
languages. I'll try to find evidence for shibboleth occurences (which
can describe words that are typical in certain registers AND provide the
potential to appear differently in phonologic realisation or semantic
expression depending on the speaker, making them behave as specific
in-group markers) in yiddish and frisian language as well as in berlin
vernacular.

\subsubsection{inspiration}\label{inspiration}

\paragraph{password parsley}\label{password-parsley}

The login to one of the protected pages of a random organisation
connected with our university is often carried out using a process known
as single sign-on. In this process, the username and password (our ZEDAT
\emph{credentials}) are not transmitted to the page we want, but only
the confirmation from an instance such that our credentials are correct,
i.e., that we have provided the instance (the ``guardian'') the correct
username and password to access the remote page (a corpus engine, a
library portal, or generally a remote application on the network that
only allows access to \emph{affiliates} of certain educational
institutions, e.g.) Both sides, guardian and repository, agree that we
are only allowed access to the resources if we are who we claim to be.
For example, a student at the FUB. Or a partisan. Or a member of a
secret society. Or: a confidant of secrets. More precisely: a
connoisseur of the SHIBBOLETH.

\subsubsection{evidence}\label{evidence}

uljana wolf, etymologischer gossip: petersilie: one will no longer be
able to use the word unreflected after reading about the 20,000 Haitian
guest workers (cf. Wolf
(\citeproc{ref-wolf_etymologischer_2021}{2021})). The word petersilie --
parsley -- /\hyperref[tofollow]{perejil}/ is needed as a means of access
in order NOT to appear as a stranger to the community of natives = to
fall victim to the massacre. Anyone who pronounces it {[}pɛʟɛχɪʟ{]} and
not {[}pɛʀɛχɪʟ{]}, as the locals do, will be murdered. So it is good to
know the pronunciation. Or to master it.

Something similar, though less drastic, can happen to people who, to
name just a few prominent examples, mispronounce
\hyperref[tofollow]{derrida}, \hyperref[tofollow]{bourdieu},
\hyperref[tofollow]{accessoir} or, to return to the subject,
\hyperref[tofollow]{shibboleth} (\he{שִׁבֹּלֶת}, cf. Sefaria
(\citeproc{ref-sefaria_judges_2025}{2025})) or, to begin the queries,
say {[}balkoːɴ{]} instead of {[}balkɔɳ{]} and /tram/ instead of
/straszenbahn/.

\subsubsection{methods}\label{methods}

Starting from a GPT provided bibliography (disclaimer:of which some
entries appear already familiar\ldots{} keywords: shibboleth, group
identity, sociolinguistics), I'll dive into corpora provided and intend
to consult speakers for their impressions. That may grow to a
quantitative corpus based study or either be limited to a qualitative
field investigation. I will not only search evidence in natural language
but also in literature, which may be easier accessible.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-arhammar_beitrage_2000}
Århammar, Nils. 2000. \emph{Beiträge Zur Nordfriesischen {Philologie}}.
Nordfriisk Instituut.

\bibitem[\citeproctext]{ref-bailey_dialect_2002}
Bailey, Charles-James. 2002. {``Dialect {Recognition}, {Group}
{Boundaries}, and {Shibboleths}.''} \emph{American Speech}.

\bibitem[\citeproctext]{ref-feitsma_saterfrisian_2010}
Feitsma, Anne. 2010. {``Saterfrisian Sociolinguistic Situation.''}
\emph{Nordic Journal of Linguistics}.

\bibitem[\citeproctext]{ref-gorter_new_1994}
Gorter, Durk. 1994. {``A New Sociolinguistic Survey of the {Frisian}
Language Situation.''} \emph{Journal of Multilingual and Multicultural
Development}. \url{https://doi.org/10.1080/03096564.1994.11784030}.

\bibitem[\citeproctext]{ref-heeringa_dialect_2001}
Heeringa, Wilbert, and John Nerbonne. 2001. {``Dialect Distance and
Speaker Perception: {The} Case of {Frisian}.''} In \emph{Computer
{Methods} in {Dialectometry}}.

\bibitem[\citeproctext]{ref-heyen_digitale_2021}
Heyen, H. 2021. {``Digitale Nordfriesische {Kommunikation}.
\#Hokerbeest?''}

\bibitem[\citeproctext]{ref-hoekstra_frisian_2022}
Hoekstra, Jarich. 2022. {``Frisian {Shibboleths}: {Phonological} and
{Lexical} {Markers} of {Group} {Identity}.''} \emph{Us Wurk}.

\bibitem[\citeproctext]{ref-katz_shibboleth_2004}
Katz, Dovid. 2004. {``Shibboleth {Phenomena} in {Eastern} {Yiddish}:
{Identity} and {Group} {Boundaries}.''} \emph{International Journal of
the Sociology of Language}.

\bibitem[\citeproctext]{ref-kvist_prosodic_2019}
Kvist, M. R. 2019. {``Prosodic {Features} of {North} {Frisian}
{Dialects}.''} \{PhD\} \{Thesis\}, University of Copenhagen.

\bibitem[\citeproctext]{ref-lonn_shibboleths_2010}
Lönn, Michael. 2010. {``Shibboleths as {Social} {Indexes} in {Germanic}
{Languages}.''} \emph{Journal of Germanic Linguistics}.

\bibitem[\citeproctext]{ref-muller_berlin_2018}
Müller, Andrea. 2018. {``Berlin {Dialect} as {Shibboleth}: {Indexical}
{Features} in {Urban} {German}.''} \emph{Zeitschrift Für Dialektologie
Und Linguistik}.

\bibitem[\citeproctext]{ref-niebuhr_pointed_2015}
Niebuhr, Oliver, and Jarich Hoekstra. 2015. {``Pointed and
Plateau-Shaped Pitch Accents in {North} {Frisian} ({Fering}).''}
\emph{Linguistische Berichte}.
\url{https://doi.org/10.1515/lp-2015-0013}.

\bibitem[\citeproctext]{ref-sefaria_judges_2025}
Sefaria. 2025. {``Judges 12.6. {Sefaria}: A {Living} {Library} of
{Jewish} {Texts} {Online}.''}
\url{https://www.sefaria.org/Judges.12.6?lang=bi&with=all&lang2=en}.

\bibitem[\citeproctext]{ref-trudgill_shibboleths_2006}
Trudgill, Peter. 2006. {``Shibboleths and {Social} {Meaning}.''} In
\emph{Sociolinguistic {Variation} and {Identity}}. Oxford University
Press.

\bibitem[\citeproctext]{ref-voeten_listener_2024}
Voeten, Chris, Anne-Fleur Pinget, Markus Kingma, Nora Stefan, and Hans
Van de Velde. 2024. {``Listener Factors in Accent Recognition: {A}
Perceptual-Dialectology Study of {Frisian}.''}

\bibitem[\citeproctext]{ref-wolf_etymologischer_2021}
Wolf, Uljana. 2021. \emph{Etymologischer {Gossip}: {Essays} Und {Reden}
/ {Uljana} {Wolf}.} 1. Auflage. Kookbooks {Reihe} {Essay} 7. Berlin:
kookbooks.

\end{CSLReferences}




\end{document}
