\documentclass[]{tufte-handout}

% ams
\usepackage{amssymb,amsmath}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \makeatletter
  \@ifpackageloaded{fontspec}{}{\usepackage{fontspec}}
  \makeatother
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \makeatletter
  \@ifpackageloaded{soul}{
     \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
     \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
   }{}
  \makeatother

\fi

% graphix
\usepackage{graphicx}
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}

% booktabs
\usepackage{booktabs}

% url
\usepackage{url}

% hyperref
\usepackage{hyperref}

% units.
\usepackage{units}


\setcounter{secnumdepth}{-1}

% citations
\usepackage{natbib}
\bibliographystyle{plainnat}


% pandoc syntax highlighting

% table with pandoc

% multiplecol
\usepackage{multicol}

% strikeout
\usepackage[normalem]{ulem}

% morefloats
\usepackage{morefloats}


% tightlist macro required by pandoc >= 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% title / author / date
\title[xtitle]{xtitle: proposition \& coherence observations in
:schizophrenia: threads}
\author{st. schwarz}
\date{2025-06-30 07:59:58.299343}


\begin{document}

\maketitle




\begin{verbatim}
## [1] "/Users/guhl/Documents/GitHub/SPUND-LX/psych/HA/"
\end{verbatim}

\begin{verbatim}
## [1] 9
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{xTitle}\label{xtitle}

\section{proposition \& coherence in :schizophrenia:
threads}\label{proposition-coherence-in-schizophrenia-threads}

\subsubsection{stephan schwarz / a. stefanowitsch:16827\_25S:sprache und
psychose}\label{stephan-schwarz-a.-stefanowitsch16827_25ssprache-und-psychose}

\subsection{subject}\label{subject}

Reference marking, coherence and information structure in schizophrenia
language.\footnote{snc.1:h2.pb.1000char/pg.queries} \#\# background
Inspired by Zimmerer et alii (\#REF) we are interested in observations
concerning coherence and propositional conditions in schizophrenia
language, as these linguistic markers appear underinvestigated in
research while they seem to play a crucial role within target group
language. (As such seen as asset of thinking or world building capacity
which might suffer from linguistic deficits within the range of positive
symptoms.) \#\# method To compute distances we queried a corpus for
matching conditions where certain (assumed) determiners appear before
similar nouns. This distance should give us information structural
evidence of how strong these noun occurences are connected, i.e.~if a
noun appears out of the blue mostly or if it somewhere before has been
introduced to the audience. In information structure definitions this
would be termed with \textbf{given and new information} Prince
(1981\#REF). ---- \#\# questions Measuring the referent-reference
distance which we here assume as indicator of coherence we hope to find
empirical evidence for disturbed or not world building capabilities
within schizophrenia language. \#\# daten We built a corpus of the
reddit r/schizophrenia thread (\texttt{n=747089} tokens) and a reference
corpus of r/unpopularopinion (\texttt{n=265670}). The corpus has been
pos-tagged using the R udpipe:: package \#REF which tags according to
the universal dependencies tagset maintained by \#REF. Still the 747089
tokens can only, with the workflow of growing the corpus and devising
the noun distances developed be just a starting point from where with
more datapoints statistical evaluation becomes relevant first. ----

\subsection{results}\label{results}

\begin{verbatim}
##    q dist  range corp corp_size      m   m_rel
## 1  a   36  446.0  obs    747089 747089 1.00000
## 2  b   70 1451.0  obs    747089  11415 0.01528
## 3  c   49  807.0  obs    747089  12516 0.01675
## 4  d   49  834.0  obs    747089  15141 0.02027
## 5  e   53  917.0  obs    747089   6983 0.00935
## 6  f   57 1119.0  obs    747089   4236 0.00567
## 7  a   40 1619.5  ref    265670 265670 1.00000
## 8  b   41 2140.0  ref    265670   4213 0.01586
## 9  c   60 2116.5  ref    265670   6542 0.02462
## 10 d   51 1863.0  ref    265670   6349 0.02390
## 11 e   63 2947.5  ref    265670    662 0.00249
## 12 f   46 2473.5  ref    265670   1576 0.00593
\end{verbatim}

\subsection{}\label{section}

\subsection{conclusion}\label{conclusion}

\subsection{B. REF:}\label{b.-ref}

\bibliography{FRZ.bib}



\end{document}
