# methods

please cf. @schwarz_this_2026 [#todo] for the corpus building and evaluation scripts

## data

```{r,m1}
#| echo: false
#| output: false
#| eval: false

mh<-lmdf.c$target=="0-human"
mp<-lmdf.c$target=="post"
mg<-lmdf.c$target=="gpt"

sh<-sum(lmdf.c$freq[c(which(mh))])
sp<-sum(lmdf.c$freq[c(which(mp))])
sg<-sum(lmdf.c$freq[c(which(mg))])
ts<-data.frame(target=c("gemini","human-pre","human-post"),tokens=c(sg,sh,sp))

```

our corpus consists of interview transcripts available in raw text which were tokenised and segmented into 10-unit chunks for further processing, @sec-clip-01.

### corpus stats
`knitr::kable(ts)` [insert corpus stats]

## clip score computation {#sec-clip-01}
the computation of clip scores was done using HU resources with a python script cf. @nenchev_clip_2026
```{r,prompt}
#| echo: false
#| warning: false
#| eval: false

ptx<-readLines("../g-prompt01.txt")
print(ptx)
```


# evaluation
## basic descriptive
## linear regression {#sec-lm-1}
to prove descriptive results, we compute the stability [...] with a linear regression model using R's lme4::lmer() function, cf. @bates_fitting_2015. coefficents are printed below [...]

```{r,lm-models}
#| echo: false
#| output: false
#| warning: false
#| eval: false

library(lme4)
library(lmerTest)

lm1<-lm(f.rel~target*in.gp,lmdf.c)
lm2<-lmer(f.rel~target*in.gp+(1|lemma),lmdf.c)

```

### basic (lm)
formula: `frequency.relative ~ target * in.gpt`

```{r,lm}
#| echo: false
#| output: true
#| warning: false
#| eval: false

summary(lm1)

```

### mixed effects model (lmer)
formula: `frequency.relative ~ target * in.gpt +(1|lemma)`

```{r,lmer}
#| echo: false
#| output: true
#| warning: false
#| eval: false

s2<-summary(lm2)
s2
ingpvar<-s2$vcov@factors$correlation
ingpv<-ingpvar[,5]
ingpv<-ingpv[4]
#ingpv
pc8<-round(s2$coefficients[8,1],6)
pc1<-round(s2$coefficients[1,1],6)
#pc1<-s2$coefficients[1,1]
pcp<-round(pc8/pc1*100,2)
pc8<-round(s2$coefficients[8,1],6)
pc1<-round(s2$coefficients[1,1],6)
pcd<-pc1-pc8
pcd<-round(pcd,5)
#pcd
p100<-pc1*100
pcdp<-(pcd/p100)
#pcdp
pcdp<-round(pcdp,5)
pcd<-pc8
p100<-100/pc1
pcdp<-(pcd*p100)
#pcdp
pcdp<-round(pcdp,5)
#pcdp

#pcdp
#pcdp
#pc1<-s2$coefficients[1,1]
pcp<-round(pc8/pc1*100,2)

```

#### helper interpretation, to be tested
the coefficients interesting for us are 

# references {#references}