# play along model collapse
## Qs
1. (script) > @schwarz_model_2026
2. (corpus) > @glaser_benjaminfeldkraft_2024

## method
the model (not pretrained cvd. blank) has been trained (G1) on above corpus with above linked script (gpt proposal using @falbel_torch_2026 in [R]) and retrained (G2) on the same corpus plus 0.1% synthetic output of G1. 

generally the training data (200.000 words) is not enough to output serious content even in G1, but we see literally that the model degenerates quickly; we observe entropy drop and decreasing vocabulary (89 vs. 86 char types).

## output G1, M3
{{< include synthetic_gen_3-1.md >}}

## output G2, M3
{{< include synthetic_gen_3-2.md >}}

## output M7d
> showing another output of G1, this time the model has been trained on token=word level, which results in far more readable, nevertheless crumbled text... / <del>the G2 ff. of that model is still to test after script adaptation.</del> generation 2/3 below.

### G1
{{< include synthetic_gen_7c-1.md >}}

### G2
{{< include ../test/modeling/synthetic_gen_7d-2.md >}}

### G3
{{< include ../test/modeling/synthetic_gen_7d-3.md >}}

### model stats
```{r, echo=F, warning=F}
#| fig-cap: "M7d model settings and statistics"
#| label: model-stats

dfa<-read.csv("../test/modeling/collapse_metrics-7d-3.csv")
knitr::kable(dfa)
  #plot(dfa$entropy,type="l",main="entropy drop over generations",xlab="generation",ylab="entropy")

```

```{r, echo=F, warning=F}
#| fig-cap: "M7d entropy develoment"
#| label: entropy-plot

# dfa<-read.csv("../test/modeling/collapse_metrics-7d-3.csv")
#knitr::kable(dfa)
  plot(dfa$entropy,type="l",main="entropy drop over generations",xlab="generation",ylab="entropy")

```

we observe decreasing entropy factors from G1 forward which means fewer diversity, higher predictabilty and a less balanced synthesis of generated texts. (cf. @leipzig_corpora_collection_korpus_2024; Entropy: ~11.7 in german newspaper corpus)

