```{r, echo=F,output=F,warning=F}
knitr::read_chunk("lxtech/benjamin.R")

```
```{r prep,echo=F,eval=T,ouput=F}
3+4
```

```{r setup-plot,echo=F,eval=T,output=T,warning=F}
```

#### this is 
just the distribution of nouns, verbs, adverbs, adjectives across the corpus plotted. 

witch corpus? > @glaser_benjaminfeldkraft_2024

```{r v2img,echo=F,eval=T,output=T,warning=F,fig.cap="corpus part of speech (nouns, verbs, adj) vector plot"}
```

#### this is 
the corpus [first 1000 words] nouns (blue) vs verbs (red) plotted 
```{r v2strips,echo=F,eval=T,output=T,warning=F,fig.cap="corpus NOUNS vs. VERBS plot"}
```

#### so you can 
create yourself a beautiful taschentuch or tischdecke or boxershort of your response paper :)

or a QR code of nouns...
```{r v2qr,echo=F,eval=T,output=T,warning=F,fig.cap="corpus[1:300] NOUNS plot"}
```

#### or any
pattern from whatever you find interesting (beautiful) to plot:

e.g.

- ich vs. du
- haben vs. sein
- past vs. present
- schund vs. high language (this may be difficult)
- you can as well plot 1 vs. n dimensions of features witch would result in a more colorful tweed