---
title: glossar
subtitle: "term global glossary"
# format: html
toc: true
toc-depth: 3
toc-expand: true

# tmeta: $testmeta$
# date: '`r format(Sys.time())`'
date: "2026-05-30"
title-meta: ""
bibliography: 
   - litKI.bib
   - LXtech.bib
   - LFG.bib
   - textur.bib
   - nietzsche.bib
   - VSstr.bib
cmeta: |
  common: global term glossary
  term: SS26
  snc: 16232.1
#[<< back to online paper <<](https://esteeschwarz.github.io/DC-LX/essais/papers/010/)  
#  arkiv: [<< the anarkkiv lankalevikki worksheet collection >>](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
arkiv: 16162.tiny
# foot: the anarkkiv lankalevikki worksheet collection >> <https://ada-sub.dh-index.org/blog>
cfoot: |
  [duumy foot collection](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
---

{{< meta cmeta >}}

{{< include _gitcom.qmd >}} 

-----

> [index](index.qmd) / glossary

-----

```{r,echo=F,eval=T}
print("engine on...")
```

```{r,echo=F,eval=T,output=F}
### get vocab, keywords
print(getwd())
f<-list.files(paste0(Sys.getenv("GIT_TOP"),"/SPUND-LX/play/quarto/start/"),pattern=".qmd",full.names = T)
f
v1<-lapply(f,function(x){
  t<-readLines(x)
  v1<-unlist(strsplit(t," "))
  m<-grepl("[^a-zA-Z]",v1)
  v1<-v1[!m]
})
v2<-unlist(v1)
head(v2)
t1<-table(v2)
#############
### rel. freq
t2<-lapply(v1,function(x){
  table(x)
})
print(t2)


```


# index
> samling, all TODO

## A
### alphabet
a common way of segmenting

### apoll
#### apollinisch
### asynchron

## D
### diachron
### dionys
#### dionysisch

## E
### embeddings
### entropie

## H
### hermeneutik
#### hermeneutischer zirkel
### hyperlink
#### related
- [intertext](#intertext)

## I
### intertext

## M
### model
#### related
- [embeddings](#embeddings), [vektor](#vektor), [training](#training)

## S
### synkretismus

## T
### tautologie
something explaining itself by itself

### training

## V
### vektor

## Z
### zipfs law