---
title: "session notes"
subtitle: "Einführung in die Grammatiktheorie und in die computationelle Modellierung am Beispiel der Lexikalisch-Funktionalen Grammatik (LFG)"

format: html
date: "2026-04-28"
toc: true
toc-depth: 3
toc-expand: true

---

::: {.content-visible when-format="html"}

-----

> [LFG](lfg-index.qmd) / session notes

-----
:::

# task session

# 1. Übung, session 3
Geben Sie die

1. Lexikoneinträge aller Worteinheiten [Tipp: überlegen Sie sich
parallel dazu, wie ein nicht-annotierter Phrasenstrukturbaum
aussieht.]
2. annotierte Phrasenstrukturregeln in der c-Struktur
3. f- inkl. a-Struktur

für den folgenden Satz des Englischen wieder:

```{r,echo=F,eval=T,warning=F}
print("The striker shot an incredible goal.")
df<-read.csv("ip1.csv")
knitr::kable(df)
```


### lexikon

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon raw"
#| label: cdf-echo
print("The striker shot an incredible goal.")

df<-readLines("ip2.csv")
print(df)
```

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon"
#| label: cdf

df<-read.csv("ip2.csv")
knitr::kable(df)
```

### ann. c-struktur

### notes
it seems as if the LFG conventions are preparing such that can be described with vectors at the end. in the way that we annotate a token (lexical item) according to specified (necessary per investigated language) categories (a,c,f - levels) which may be projected into a vector space. say if we have 10 categories with key-value pairs of each e.g. 5 attributes (keys) which can each again take 5 values which will then be coded TRUE or FALSE , we will have the token represented in a vectorspace of `10x5x5x2` = ```r 10*5*5*2```  dimensions, nestcepas? well thats a lot..., but seeing that model vectorspaces use to have embeddings dimensions in that range (e.g. 762 for stable diffusion models) its not strange assuming that these amount of categories and key-value pairs is enough...

## process
first the categories (the ones from above) have to be coded binary. so we have:

- the syntax head class (leX), which in our example take:
  - N
  - A
  - D
  - V
- PRED
  - subj
  - obj
- NUM
  - sg
  - pl
- PERS
  - 1
  - 2
  - 3
- DEF
  - plus
  - minus

`striker` coded binary would look like: `X.N=1,NUM.sg=1`. all other categories which are not 1(TRUE) default to FALSE. then the complete row would be:

`t=striker,X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0`

```{r echo=F}
df<-data.frame(rbind(
c(t="the",X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,TENSE.pre=0,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=1,DEF.minus=0),
c(t="striker",X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,TENSE.pre=0,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0),
c(t="shot",X.N=0,X.A=0,X.D=0,X.V=1,PRED.subj=1,PRED.obj=1,NUM.sg=0,NUM.pl=0,TENSE.pre=0,TENSE.past=1,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0),

c(t="an",X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,TENSE.pre=0,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=1),

c(t="incredible",X.N=0,X.A=1,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,TENSE.pre=0,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0),

c(t="goal",X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,TENSE.pre=0,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0)
))
knitr::kable(df)
#df
#c1<-cos(df[1,2:length(df)])
#c2<-cos(df[2,2:length(df)])
```

### Q
question is: shall one code (with shot e.g.) the NUM category as it references sg., which is still not visible as being not marked but accessible via knowledge? and PERS as well as its also not marked (in past tense) but known 3rd p.? and: can we code DEF alone with 1/0 for plus/minus or do we need 2 cats here?

> here "context" comes into play: we assume that lexicon entries are :context free: and code only visible features.

Q3, another one: 'shot' must in the lexikon be coded also as X.N=true and NUM.sg=true as its without context not visible if its the noun or the past tense verb, nestcepas?

### play 
- cosine similarity q/vector
- categories cluster

```{r,echo=F,eval=T}
qd<-c(t="goes",X.N=0,X.A=0,X.D=0,X.V=1,PRED.subj=1,PRED.obj=0,NUM.sg=0,NUM.pl=0,TENSE.pre=1,TENSE.past=0,PERS.1=0,PERS.2=0,PERS.3=1,DEF.plus=0,DEF.minus=0)
df<-rbind(df,qd)
q<-qd[1]
knitr::read_chunk("lxtech/cosine-sim.R")
```

#### Q3
and in Q3: if we also code animate vs. inanimate and OBJ.theta.target N then the clustering of goal|striker would also look different since not both can take true arguments of active-only and obligatory target-only objects verbs - if e.g. 'shot'-X.V and 'goal'-OBJ.target...   
but: when 'shot' takes (^PRED)="shoot<sub.obj>" it - coded alone - does not show above ambiguity, saying the predicate qualifier 'shoot' here is important for the rest of the row. so further: we either have to somehow embed the qualifying reference into the binary system (which is not possible since its again an (infinite space of tokens/ lexical items[lemmas])) or distinguish tokens by token qualities like 'shot.shoot' vs. 'shot.shot'. 

::: {.panel-tabset}

#### shot original (proposed)


```{r set1,echo=F}
#m<-df[,"t"]=="shot"
#df<-df
#df[m,c(2,8)]<-c(1,1)
knitr::kable(df)
```

```{r cosine-sim,echo=F}
```

```{r plot1,echo=F}
```

#### modified "shot"

```{r set2,echo=F}
m<-df[,"t"]=="shot"
#df<-df
df[m,c(2,8)]<-c(1,1)
knitr::kable(df)
```

```{r cosine-sim,echo=F}
```

```{r plot2,echo=F}
```

#### shot disambigue
```{r set3,echo=F}
m<-df[,"t"]=="shot"
#df<-df
df[m,c(2,8)]<-c(0,0)
df[m,1]<-"shot.shoot"
dt<-df[m,]
dt$t<-"shot.shot"
dt[,c(2,5,6,7,8,11)]<-c(1,0,0,0,1,0)
df<-rbind(df,dt)
#df
knitr::kable(df)
```

```{r cosine-sim,echo=F}
```

```{r plot3,echo=F}
```

the thing here is that theres no rendering of the TENSE difference of goes|shot.shoot since height (y-axis) is only displaying raw difference in sum and they sum up equal in TRUE/FALSE.

:::

### functions, equations, constraints, f-precedence
can be we at this point arrive at the sense of the functional level. our only way to binary code unambigously and despite have all visible features if not coded so at least checked for, is to make use of functions that have a binary output. so for the shot.shoot token the function would ca. be: check, if theres a noun preceding with no verb between. if so, our token is coded as a verb. if we apply that top-down then we'd be able to devise for occuring verbs.  

### coding task
#### goal
plot output phrasenstrukturbaum from binary f-structure model via hclust()

#### method
as the binary coded dataframe already includes divergence information on nodes and ranks them according to level of divergence it should be possible to map the diverging binary values after applying a hierarchy to the keys(attributes) such that the position of the attribute within the dataframe (column) influences the height relation of nodes. 

#### disclaimer
the above displayed binary dataframe is *not* derived from script but manually created as its only visualing (as sample) what woud be the goal to achieve then via a script, such that a parser (script) would look at a sentence/token and construct the binary code without context available. this is still the main question what can be derived automatically (with a lexicon available) and what not. (a common postagger does exactly that.) my personal aim here was rather to complete the mapping on base of a binary code without further annotation necessary and that seems feasable...

## T2 NT, session 5
`meine schwester schreibt an einem wichtigen buch`

```
IP --> VP --> DP --> NP D meine N schwester 
VP --> V schreibt
IP --> PP --> P an DP --> NP --> D einem --> AdP Adv wichtigen N buch
```

```
IP --> DP
IP --> I' --> VP --> V'
V' --> V
V' --> PP --> an einem wichtigen buch
```

```
IP --> NP meine schwester
IP --> I' --> VP
VP --> V schreibt
VP --> PP
PP --> P an
PP --> NP einem wichtigen buch
```

```
IP --> NP
IP --> I --> VP --> V' --> V schreibt
V' --> PP
PP --> P an
PP --> NP --> Det einem
NP --> N' --> AP wichtigen
N' --> N buch
NP --> Det meine
NP --> N' --> schwester
```

## tree s
### snc
123456
```{r, echo=F,output=F,warning=F}
knitr::read_chunk("lxtech/trees.R")

```
```{r tree-prep,echo=F,eval=T,ouput=F,warning=F}
lines<-"IP --> NP1 --> Det meene
NP1 --> N'1 --> N1 schwester
IP --> I --> VP --> V' --> V schreibt
V' --> PP --> P an
PP --> NP2 --> Det einem
NP2 --> N'2 --> AP wichtigen
N'2 --> N2 buch"

t3<-"../output/pages/004/t4"
f<-"t3.txt"
#f
#getwd()
lines<-readLines(f)
cat("raw used final:\n")
print(lines)
f<-unlist(strsplit(f,"\\."))[1]
#f
t3<-paste0("../output/pages/004/",f)
#t3
dpdf<-F
size<-70
```


```{r m2forest,echo=F,eval=T,ouput=F,warning=F}
```

```{r outpdf,echo=F,eval=T,warning=F,results="asis",output=T}
```
