---
title: LFG
format: html
date: "2026-04-28"

---

# task session 3

## 1. Übung
Geben Sie die

1. Lexikoneinträge aller Worteinheiten [Tipp: überlegen Sie sich
parallel dazu, wie ein nicht-annotierter Phrasenstrukturbaum
aussieht.]
2. annotierte Phrasenstrukturregeln in der c-Struktur
3. f- inkl. a-Struktur

für den folgenden Satz des Englischen wieder:

```{r,echo=F,eval=T,warning=F}
print("The striker shot an incredible goal.")
df<-read.csv("ip1.csv")
knitr::kable(df)
```


### a-struktur, lexikon

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon raw"
#| label: cdf-echo
print("The striker shot an incredible goal.")

df<-readLines("ip2.csv")
print(df)
```

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon"
#| label: cdf

df<-read.csv("ip2.csv")
knitr::kable(df)
```

### ann. c-struktur

# notes
it seems as if the LFG conventions are preparing such that can be described with vectors at the end. in the way that we annotate a token (lexical item) according to specified (necessary per investigated language) categories (a,c,f - levels) which may be projected into a vector space. say if we have 10 categories with key-value pairs of each e.g. 5 attributes (keys) which can each again take 5 values which will then be coded TRUE or FALSE , we will have the token represented in a vectorspace of `10x5x5x2` = ```r 10*5*5*2```  dimensions, nestcepas? well thats a lot..., but seeing that model vectorspaces use to have embeddings dimensions in that range (e.g. 762 for stable diffusion models) its not strange assuming that these amount of categories and key-value pairs is enough...

## process
first the categories (the ones from above) have to be coded binary. so we have:

- the syntax head class (leX), which in our example take:
  - N
  - A
  - D
  - V
- PRED
  - subj
  - obj
- NUM
  - sg
  - pl
- PERS
  - 1
  - 2
  - 3
- DEF
  - plus
  - minus

`striker` coded binary would look like: `X.N=1,NUM.sg=1`. all other categories which are not 1(TRUE) default to FALSE. then the complete row would be:

`t=striker,X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0`and applied to the other tokens:

`t=the,X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=1,DEF.minus=0`

`t=shot,X.N=0,X.A=0,X.D=0,X.V=1,PRED.subj=1,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=1,DEF.plus=0,DEF.minus=0`

`t=an,X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=1`

`t=incredible,X.N=0,X.A=1,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0`

`t=goal,X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0`


```{r echo=F}
df<-data.frame(rbind(
c(t="the",X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=1,DEF.minus=0),
c(t="striker",X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0),
c(t="shot",X.N=0,X.A=0,X.D=0,X.V=1,PRED.subj=1,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=1,DEF.plus=0,DEF.minus=0),

c(t="an",X.N=0,X.A=0,X.D=1,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=1),

c(t="incredible",X.N=0,X.A=1,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=0,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0),

c(t="goal",X.N=1,X.A=0,X.D=0,X.V=0,PRED.subj=0,PRED.obj=0,NUM.sg=1,NUM.pl=0,PERS.1=0,PERS.2=0,PERS.3=0,DEF.plus=0,DEF.minus=0)
))
knitr::kable(df)
#df
#c1<-cos(df[1,2:length(df)])
#c2<-cos(df[2,2:length(df)])
```

