---
title: LFG
format: html
date: "2026-04-28"

---

# task session 3

## 1. Übung
Geben Sie die

1. Lexikoneinträge aller Worteinheiten [Tipp: überlegen Sie sich
parallel dazu, wie ein nicht-annotierter Phrasenstrukturbaum
aussieht.]
2. annotierte Phrasenstrukturregeln in der c-Struktur
3. f- inkl. a-Struktur

für den folgenden Satz des Englischen wieder:

```{r,echo=F,eval=T,warning=F}
print("The striker shot an incredible goal.")
df<-read.csv("ip1.csv")
knitr::kable(df)
```


### a-struktur, lexikon

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon raw"
#| label: cdf-echo
print("The striker shot an incredible goal.")

df<-readLines("ip2.csv")
print(df)
```

```{r,echo=F,eval=T,warning=F}
#| fig-cap: "lexikon"
#| label: cdf

df<-read.csv("ip2.csv")
knitr::kable(df)
```

### ann. c-struktur

# notes
it seems as if the LFG conventions are preparing such that can be described with vectors at the end. in the way that we annotate a token (lexical item) according to specified (necessary per investigated language) categories (a,c,f - levels) which may be projected into a vector space. say if we have 10 categories with key-value pairs of each e.g. 5 attributes (keys) which can each again take 5 values which will then be coded TRUE or FALSE , we will have the token represented in a vectorspace of `10x5x5x2` = ```r 10*5*5*2```  dimensions, nestcepas? well thats a lot..., but seeing that model vectorspaces use to have embeddings dimensions in that range (e.g. 762 for stable diffusion models) its not strange assuming that these amount of categories and key-value pairs is enough...