---
title: "LX / Tech"
subtitle: "Sprache und Technologie - Language and Technology"
format: html
# tmeta: $testmeta$
date: '`r format(Sys.time())`'
title-meta: ""
categories:
   - vectors
   - noun compounds
bibliography: LXtech.bib
nocite: '@*'
cmeta: |
  class: [16830] Sprache und Technologie - Language and Technology::
  tutor: frankowsky ::
  term: SS26
  snc: 16175.1   
#[<< back to online paper <<](https://esteeschwarz.github.io/DC-LX/essais/papers/010/)  
#  arkiv: [<< the anarkkiv lankalevikki worksheet collection >>](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
arkiv: 16162.tiny
# foot: the anarkkiv lankalevikki worksheet collection >> <https://ada-sub.dh-index.org/blog>
cfoot: |
  [duumy foot collection](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
---

::: {.content-visible when-format="html"}
{{< meta cmeta >}}
:::

{{< include _gitcom.qmd >}} 

```{r,echo=F,eval=T}
print("engine on...")
#val <- quarto::quarto_metadata$get("meta")
#meta <- knitr::metadata
#params <- meta$
#val<-params$tmeta
val.f<-"fixed val"
val.c<-"{{< meta testmeta >}}"
val.c<-"{{< meta cmeta >}}"
```

#### snc
- button.12345

# index
1. samling: margin notes general > [open](lxtech-qa.qmd){.btn .btn-cat} 


## samling
- @frankowsky_extravagant_2022
  - how were dates attested technically?
  - what about the decow16 corpus?
    - possible to put that online (term paper) or is it? (@schafer_what_2025)
    - corpus formats?

## zinsmeister
### play
```{r,echo=F,eval=T}
knitr::read_chunk("lxtech/skewdiv.R")
```

```{r skew}
```

### notes
the tokens vector sujet gets interesting if following [this](lfg-001.qmd) theories (LFG: lexical functional grammar) where we may describe token categories in a vector space as well, undisputed of distribution. so if @zinsmeister_corpusbased_2013 investigate on similarity based on distribution of lexical items in a context window (coocurrences) we may also (like in LLM embeddings) make statements about token relatedness based on the vector cosine similarity of tokens semantic / syntactic categories. 