---
title: "LX / Tech"
subtitle: "Sprache und Technologie - Language and Technology"
format: html
toc: true
toc-depth: 3
toc-expand: true
# tmeta: $testmeta$
date: '`r format(Sys.time())`'
title-meta: ""
categories:
   - vectors
   - noun compounds
   - llms
bibliography: LXtech.bib
nocite: '@*'
cmeta: |
  class: [16830] Sprache und Technologie - Language and Technology::
  tutor: frankowsky ::
  term: SS26
  snc: 16175.1   
#[<< back to online paper <<](https://esteeschwarz.github.io/DC-LX/essais/papers/010/)  
#  arkiv: [<< the anarkkiv lankalevikki worksheet collection >>](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
arkiv: 16162.tiny
# foot: the anarkkiv lankalevikki worksheet collection >> <https://ada-sub.dh-index.org/blog>
cfoot: |
  [duumy foot collection](https://stschwarz.userpage.fu-berlin.de/links/?title=paperfoot&ref=paper-temp)
---

::: {.content-visible when-format="html"}
{{< meta cmeta >}}
:::

{{< include _gitcom.qmd >}} 

-----

> [index](index.qmd) / LXTech 

-----

```{r,echo=F,eval=T}
print("engine on...")
#val <- quarto::quarto_metadata$get("meta")
#meta <- knitr::metadata
#params <- meta$
#val<-params$tmeta
val.f<-"fixed val"
val.c<-"{{< meta testmeta >}}"
val.c<-"{{< meta cmeta >}}"
```

#### snc
- button.12345678
- chunk out

# index
1. samling: margin notes general > [open](lxtech-qa.qmd){.btn .btn-cat} 
2. LLM and LX theories, play claude > [open](lxtech003.qmd), original: [claude](https://claude.ai/share/404f1ff7-0f1a-48b7-b449-bdf255996f5e)

## samling
- @frankowsky_extravagant_2022
  - how were dates attested technically?
  - what about the decow16 corpus?
    - possible to put that online (term paper) or is it? (@schafer_what_2025)
    - corpus formats?

## zinsmeister
### play
```{r,echo=F,eval=T}
knitr::read_chunk("lxtech/skewdiv.R")
```

::: {.panel-tabset}

#### 1. skew divergence
```{r skew}
```

#### 2. cosine similarity of vectors
```{r cosine-sim,echo=T,output=T}

```

#### 3. clustering embeddings

```{r prob1,echo=T,output=T}

```

:::

### notes
the tokens vector sujet gets interesting if following [this](lfg-001.qmd) theories (LFG: lexical functional grammar) where we may describe token categories in a vector space as well, undisputed of distribution. so if @zinsmeister_corpusbased_2013 investigate on similarity based on distribution of lexical items in a context window (coocurrences) we may also (like in LLM embeddings) make statements about token relatedness based on the vector cosine similarity of tokens semantic / syntactic categories.

the skew divergence approach to measure similiarities of their noun-noun compounds (which nevertheless can be applied to any type of comparison between two probabilty vectors = embeddings) applied by zinsmeister is farely easy to implement in studies since it doesnt afford much manual annotation besides (in their study) to define the contextual words for which a cooccurence is evaluated. playing with the width of a context window, the corpus base or other parameters that constrain cooccurence can yield interesting results in semantic distance of items. 

besides, one could also include other parameters next to distributional, including syntactical information, to enlarge the dimension of the vector space and overall resolution of comparison.

## LLMs
i did think about consequences of future LLMs training data will most likely include

1. recent AI discourse subjects (papers on, discussions, public discurse)
2. recent generated content itself

so i asked gpt about Q2. [her answer.](lxtech002.qmd)

### small model training G1 synthetic output

> [open](synthetic_gen_3.qmd){.btn .btn-cat}
<!-- <iframe src="synthetic_gen_3-1.html" width="100%" height="500px"></iframe> -->

# references