---
title: "interaction-copresence (replic)"
author: "esteeschwarz"
date: "`r Sys.time()`"
output: html_document
bibliography: imo-replic.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# replication study
## preliminaries
Q1: [@imo_digital_2024]. Digital vermittelte Interaktion ohne Kopräsenz: Immer noch Interaktion? In Handbuch Sprache und Digitale Kommunikation (Bd. 23, S. 249–268). Walter de Gruyter GmbH. <https://doi.org/10.1515/9783110744163-012>   

## goal
- replicate study using another corpus
- turns out to be reddit-commentaires since the first tried chat corpora either (Q2) doesnt really support export of query results or, in case of the swiss whatsapp corpus (Q3-5) doesnt support export of results of complex queries. (I am working to handle this.)

## method

```{r}
query<-"okay|\\boki\\b|\\boke\\b|\\bokey|\\bokee|okidoki|o\\.k\\.|\\bokk\\b|\\bokkk\\b"
limit<-"That|that|That's|that's|that\031s|I'm|It's|it's|Its|its"

```
----

Q2: <https://www.english-corpora.org/glowbe/>   
> no export possible of results   

### ANNIS queries, swiss whatsapp corpus:
Q3: <https://corpora.linguistik.uzh.ch/annis/?id=4d1db244-fe3d-4b2d-a47b-04d985f06ea0>   
> no export possible of results   

```{aql}
tok=‎/.*‎/ & 
tok=‎/.*‎/ & tok=‎/(okay|oki|oke|okey|okee|okidoki|o\.k\.|okk|okkk)‎‎/ & tok=‎/.*‎/ &
msg_id = ‎/.*‎/ &
msg_id = ‎/.*‎/ &
spk = ‎/.*‎/ &
spk = ‎/.*‎/ 
& #1 . #2 & #2 _=_ #5
& #2 . #3
& #3 . #4 & #3 _=_ #6
& #5 != #6 
& #2 _=_ #7 
& #4 _=_ #8
```

----

Q4: <https://corpora.linguistik.uzh.ch/annis/?id=ae12da87-b757-4a43-ba92-183194f43b11>   
> no export possible of results   

```{aql}
tok=‎/.*‎/ &  tok=‎/(okay|oki|oke|okey|okee|okidoki|o\.k\.|okk|okkk)‎/ &
msg_id = ‎/.*‎/ &
msg_id = ‎/.*‎/ 
& #1 . #2 
& #1 _=_ #3
& #2 _=_ #4
& #3 != #4 
```

----

Q5: <https://corpora.linguistik.uzh.ch/annis/?id=45d7c183-eab2-451e-9277-3ceda612c2f6>   

```{aql}
tok=‎/(okay|\boki\b|\boke\b|\bokey|\bokee|okidoki|o\.k\.|\bokk\b|\bokkk\b|\boki\b)‎/
```

----

### reddit comment extraction

Q6: <https://www.reddit.com/r/unpopularopinion/>   
script: <https://github.com/esteeschwarz/SPUND-LX/blob/main/intLX/corpus-essais.R>   

- extract ```r query```, ```n=112/25120```
- limit to token within 4 from turn start, ```n=15/112```
- limit to turn without ```r limit``` at turn start, ```n=11/15```

#### results
```{r }

load("data/com.ok.extract.RData")
print(com.ok.extract)

```

```{r }

load("data/com.df_wt_pre.RData")
print(com.ok.extract)

```


## NT: swiss chat corpus sample


```{r}
#load("https://raw.githubusercontent.com/esteeschwarz/SPUND-LX/intLX/data/swisschat_sample.RData")
d1.sm.g<-read.csv("https://raw.githubusercontent.com/esteeschwarz/SPUND-LX/main/intLX/data/swisschat_sample.csv")
#library(knitr)
print(d1.sm.g$x)
```

----
# B. references
