---
title: "Polysemous Verbs with Light and Concrete Senses"
subtitle: "A Corpus Linguistics Replication Study"
author: "Stephan Schwarz – FU Berlin – 2026"
format:
  html:
    css: poster.css
    theme: Cosmo
    toc: false
    self-contained: false
execute:
  echo: false
---
#### meta
class: germanische sprachen im vergleich  
tutor: Hüning / Konvicna  
term: WS25/26  

```{r background-script, echo=FALSE, results='asis'}
script_lines <- readLines("../LLM-003.R")

cat("```r\n")
cat(paste(script_lines[1:100], collapse = "\n"))
cat("\n```")
```


# Ausgangslage

We explored frequencies of verbs in light and concrete use in the
Santa Barbara Corpus of Spoken American English.

We aim to replicate Mehl (2021).

# Hypotheses

- Light senses are more frequent
- Concrete senses are more salient
- Embodied experience shapes meaning

# Methoden

We built a corpus database of SBC data available for download.

After annotating for light and concrete use, we counted frequencies.

# p3

```{r,fig1}
plot(cars)
```

# p2
```{r,fig2}
hist(rnorm(100))
```

```{r,fig3}
boxplot(rnorm(100), main = "")
```

# Conclusion
We observe a preference for the lemma make
in spoken context.
The create-light relation expectation
was not fulfilled.



# References
Gilquin (2008)
Mehl (2021)
UCSB Santa Barbara Corpus